\textls[-0]{\thetitle}

### 2023-03-23 metatheory2.tex
Finish Explaning the necessity of boldP.

% Deal with intuition

>>>>

% Deal with intuition
\noindentparagraph{The necessity of $\TyTkg{\cdot}$.} 
According to the equality rules of $\ruleref{tyeq/pk/add}$, for a fixed $\lsig$, $\TyTkg{\lsig}$ is structurally similar to $\TyLkg{\lsig}$ but we are using dependent tuple $\pair{\cdot}{\cdot}$ instead of $\LkgAdd{\cdot}{\cdot}$ to package the fields of a linkage. We have computed several examples in \cref{fig:stlc-linkage-typing}.

Thus the mystery of $\TyTkg{\cdot}$ really lies in its necessity. $\TyTkg{\cdot}$ is indispensible because \emph{we cannot do abstraction with $\TyLkg{\cdot}$ alone}, which in turn is due to the fact that an abstraction $A$ inside $\ruleref{lsig/add}$ and $\ruleref{l/add}$ is \emph{usually a dependent tuple type}.

\begin{align*}
  \lsig &\coloneq \{ T_A : \TyS{\wcode{\wsig_1}}; &&T_B : \TyS{\wcode{\wsig_2}}; &&f : \El{T_A} \to \El{T_B} \} \\
  \lsig^A &\coloneq \{T_A : \cU; &&T_B : \cU; &&f : \El{T_A} \to \El{T_B} \}
\end{align*}

We illustrate with a small example using named convention, say given two inductive type $\wsig_1$ and $\wsig_2$ and a function $f$ maps from the first to the second type. Then we will have a linkage $\goodTerm{\cdot}{\ell}{\TyLkg{\lsig}}$. Say we want to add a new field into this linkage and abstracting both extensible inductive types, then in $\ruleref{l/add}$, the chosen $A$ must abstract both inductive type to $\cU$, resulting something like $\lsig^A$. The question now is whether $A$ is a linkage type $\TyLkg{\lsig^A}$ or a dependent tuple type $\TyTkg{\lsig^A}$; and for each case we will intuitively choose the following approach to do abstraction.

\begin{align*}
  & \text{doing abstraction} \\ 
  \text{\textcolor{red}{Approach $\alpha$}}  \quad&\quad {\color{red} \TyLkg{\lsig} \longrightarrow \TyLkg{\lsig^A}} \\
  \text{Approach } \beta \quad&\quad  \TyLkg{\lsig} \longrightarrow \TyTkg{\lsig} \longrightarrow \TyTkg{\lsig^A}\\
\end{align*}

However, the seemingly easier approach $\alpha$ is impossible because linkage $\TyLkg{\lsig^A}$ is uninhabited---say we have $\ell^A : \TyLkg{\lsig^A}$, then by projection we will have an impossible term $\goodTerm{T_A : \cU, T_B: \cU}{\ell^A.f}{\El{T_A} \to \El{T_B}}$ due to the nature of overridability of linkage. To make the inconsistency clearer, by double abstraction we will have an impossible function $g \coloneq \goodTerm{\cdot}{\nlam{T_A}{\nlam{T_B}{\ell^A.f}}}{\nTyPi{T_A}{\cU}{\nTyPi{T_B}{\cU}{\El{T_A} \to \El{T_B}}}} $, i.e. a function maps from arbitrary type to another arbitrary type . We can easily derive inconsistency setting applying $\napp{\napp{g}{\top}}{\bot} : \top \to \bot$.

Thus we can only choose the approach $\beta$ instead, which justifies why we have $\TyTkg{\cdot}$ in our typing rule, as $\TyTkg{\lsig}$ is usually the first step to do abstraction.

To summarize, the reason we need dependent tuple type is because of the overridability nature of our linkages, together with the frequent usage of universe type in the context of dependent type programming. Once we want to abstract a type in a signature, we may have to use dependent tuple type as the abstraction instead of linkages. Since abstraction is usually dependent tuple type, we introduce $\TyTkg{\cdot}$ and its computation rules to make the process of doing abstraction in \TT easier.
